Effects of the body force on the evacuation dynamics

### 2019-12-15 paper.tex
puse el diagrama de venn en formato png
agregue el diagrama de venn con dos parametros

\section{\label{background}Background}

@@

\section{\label{simulations}Numerical simulations}

@@

@@

%/////////////////////////////////////////////////
\begin{thebibliography}{10}
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\expandafter\ifx\csname href\endcsname\relax
  \def\href#1#2{#2} \def\path#1{#1}\fi

@@

\end{thebibliography}
\end{document}
\endinput

>>>>

The Social Force Model (SFM) addresses two physical forces as 
\textit{essential}: the ``body force'' and the ``sliding friction''. Both are 
inspired by granular interactions and were claimed to be necessary  
for attaining the particular effects in panicking crowds \cite{helbing_2000}. 
The ``sliding friction'' actually proved to be an essential feature of the 
``faster-is-slower'' effect, although the role of the ``body force'' appears, 
at a first instance, not so clear \cite{dorso_2005,dorso_2007,dorso_2011}. \\

The existence of a ``body force'' in the context of highly dense crowds (say, 5 
to 10$\,$people/m$^2$) is a commonsense matter \cite{henein_2007,fruin_1993}. 
Researchers, however, question the numerical setting for this force in 
the SFM context \cite{lakoba_2005}. As a matter of fact, the usual setting by 
Helbing prevents the overlapping among pedestrians, but it is known to 
accomplish artificially high force levels 
\cite{helbing_2000,lakoba_2005,langston_2006,lin_2017}. The force estimates 
from the SFM appear to be remarkably higher with respect to the reported real 
life data (say, an order of magnitude). The crowd motion simulations, however, 
present quite realistic results \cite{lakoba_2005,langston_2006,dorso_2017}. 
The 
point seems to be that the SFM focuses on the ``incoordination phenomenon'' due 
to clogging, missing the ``individualistic'' perspective  of single pedestrians 
or very small groups \cite{helbing_2000,henein_2007,narain_2009}.  \\

Many researchers realized that modifying the SFM may (partially) surpass the 
misleading parameter setting. It was proposed that the pedestrians' 
psychological force (say, the ``social force'') should be suppressed in the 
context of highly dense crowds 
\cite{pelechano_2007,moussaid_2011,alonso_2014,bottinelli_2017}, or smoothly 
quenched according to the crowd density \cite{song_2019}. The authors in 
Refs.~\cite{kabalan_2017,jebrane_2019} further proposed a rigid body model in 
order to completely avoid the overlapping phenomenon. This perspective 
dismisses 
any connection to a ``sliding friction''. Conversely, other authors tried to 
limit the pedestrians acceleration by introducing ``static friction'' between 
the pedestrians and the floor \cite{wang_2019}. This kind of friction, however, 
reduces the effective willings of the pedestrians.  \\

A meaningful (individualistic) numerical setting appears not available yet (to 
our knowledge). The reason is that different numerical settings can lead to 
the same crowd dynamics. Actually, only a small set of dimensionless 
``numbers'' control the crowd dynamics \cite{dorso_2019}. These are similar to 
those encountered in other active matter systems (P\'eclet number, etc.)  
\cite{marchetti_2014}. We may hypothesize that while the dimensionless 
``numbers'' provide some kind of control on the ``incoordination phenomenon'' 
in 
crowds, only a few numerical settings can attain an ``individualistic'' 
meaning. 
 \\

The numerical setting for the ``faster is slower'' effect presented by 
Helbing and co-workers is somewhat cumbersome 
Ref.~\cite{helbing_2000,dorso_2017,dorso_2019}. Although a single 
parameter (say, the desired velocity $v_d$) is numerically varied to 
explain the phenomenon, the researcher looses sight of the dimensionless 
``numbers'' that truly control the crowd dynamics. The setting in 
Ref.~\cite{helbing_2000} also misses the ``faster is faster'' effect reported 
to occur at very high pedestrian densities \cite{dorso_2017,haghani_2019}. 
Alternatively, the empirical fundamental diagram raises as a point of reference 
for the SFM control parameters \cite{helbing_2007,dorso_2017}. \\

The fundamental diagram exhibits the flux behavior for either low dense crowds 
(with seldom contacts between pedestrians) and highly dense crowds (dominated 
by 
body forces and sliding friction). The latter usually experiences a flux 
slowing 
down, but other behaviors are also possible \cite{helbing_2007,lohner_2018}. In 
light of our previous hypothesis, we may suspect that the  modeling of the 
``flux slowing down'' within the context of the SFM will require the proper 
setting of the (dimensionless) controlling parameters. We examined this working 
hypothesis in Ref.~\cite{dorso_2019}, but limiting the parameter exploration to 
the sliding friction, disregarding the body force. \\

We now widen the investigation on the parameter settings to include the one 
associated to the body force. We will focus on the complex interplay between 
the body force and the sliding friction among pedestrians. Recall that the 
interplay dynamics is not directly controlled by the numerical setting, but 
through dimensionless ``numbers'', where the model parameters appear mixed 
between each other. Thus, this step up offers a challenge to the 
``individualistic'' meaning of the parameter's setting.

The investigation is organized as follows. We first recall the available 
experimental values on the body force and the sliding friction (see Section 
\ref{experimental}). Secondly, we introduce the reduced-in-units SFM and the 
three dimensionless numbers that control the crowd dynamics (see Section 
\ref{background}). We present our numerical simulations in Section 
\ref{results}. For the sake of clarity, this Section is 
separated into two major parts: the bottleneck scenario and the corridor 
scenario in Sections \ref{bottleneck} and \ref{corridor}, respectively. 
Section \ref{discussion} opens a detailed discussion raising from of the results 
in Section \ref{results}. The last Section closes the investigation with our 
main conclusions.

\section{\label{experimental}Experimental background}

\section{\label{background}Theoretical background}

@@

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.40\columnwidth]{./venn_2parameters.png}\label{venn_2parameters}}\ 
    \subfloat[]{\includegraphics[width=0.40\columnwidth]{./venn_3parameters.png}\label{venn_3parameters}}\\
\caption[width=0.47\columnwidth]{}
\label{flow_density}
\end{figure}

\section{\label{discussion}Discussion}

@@

%/////////////////////////////////////////////////
%\begin{thebibliography}{10}
%\expandafter\ifx\csname url\endcsname\relax
%  \def\url#1{\texttt{#1}}\fi
%\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
%\expandafter\ifx\csname href\endcsname\relax
%  \def\href#1#2{#2} \def\path#1{#1}\fi

@@

%\end{thebibliography}

\bibliographystyle{unsrt}
\bibliography{paper}% Produces the bibliography via BibTeX.

\end{document}
\endinput
